The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-10-28 memos/introduction_v0.tex
memos: add draft introductions

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{comment}

\shortauthors{LGB, KM, JNW}
\shorttitle{Binary Biases \& Toy Transit Surveys}

\begin{document}

% ------------------------------------------------------------------------
% New commands
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\lsim{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\gsim{\lower.5ex\hbox{\gtsima}}
\def\tess{{\it TESS} }
\def \teff {T_{\rm eff}}
\def \phir {\Phi_{\rm R}}
\def \fov {24$^{\circ}$}
\def \pixsz {21.1''}
\def \aeff {69.1 cm$^2$ }    
\def \epd {105 mm}

\def \kepler {{\it Kepler}}

% -------------------------------------------------------------------------
%

\bibliographystyle{apj}

\title{ The Effects of Stellar Binarity on Transit Survey Occurrence 
Rates }

\author{
  LGB, KM, JNW
}

% \journalinfo{Draft version}
\slugcomment{Draft version}

%\altaffiltext{1}{Princeton University}

\begin{abstract}

What errors does ignoring binarity introduce on the occurrence rates derived 
from transit surveys?

\begin{comment}
First, we show that in a universe where all stars have fixed properties
(except that some are twin binaries), the results of a search for planets of a 
given radius and orbital period can be expressed analytically.

We then allow the properties of the secondary in binary systems to vary. A 
distribution of light ratios introduces a detection bias towards 
equal-mass binaries in magnitude limited samples.
We show that dependent upon planet occurrence's functional dependence on host 
star mass, as well as the specifics of the assumed survey completeness, the 
occurrence rate correction factor $X_\Gamma \equiv \Gamma_{\rm 
true}/\Gamma_{\rm obs}$ ranges from $0.6-1.6$.

Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
show that if we assume every star in this {\it Kepler} survey analog is single, 
we underestimate occurrence rates for a given planet radius and period by 
a factor of $\approx 1.4$, or a percentage error relative to the true 
occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
true}=-28\%$. This is relevant for measuring accurate values of the rate at 
which Earth-like planets exist around Sun-like stars.
\end{comment}

\end{abstract}

\keywords{planets and satellites:\ detection}

@@

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary bounds orbiting the stars in a given volume 
of stellar phase-space $(\Delta R_p, \Delta P, \Delta M_\star,\ldots$).
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that seem to be of the desired type, 
and searchable for the planet.  These ``searchable stars'' are the points on 
the sky for which they think planets are observed with 100\% detection 
efficiency. Correcting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

\begin{equation}
\Gamma_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
\end{equation}

@@

Here we focus on problems that arise from the fact that stars of the desired
type often exist in binaries. 
We assume for simplicity that all binaries in the survey are 
spatially unresolved.

An immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets may differ 
between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval of say $\Delta 
M_\star$, it must implicitly marginalize over stellar multiplicity.
Otherwise, one must address ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

Outside of astrophysical differences, there are observational biases.
A given apparently-searchable star may truly be a single star of the desired 
type. If not, and ignoring higher order multiples, one of the following must 
be true:
\begin{enumerate}	
\item The apparently-searchable star is a binary system, with two searchable 
stars of the desired type. $N_\star$ is under-counted by one for every such 
system.
%
\item The apparently-searchable star is a binary system, with one searchable 
star of the desired type. 
The searchable star of the desired type could be the primary or the secondary.
$N_\star$ is correctly counted for every such system, though there may be 
systematic errors in estimates of stellar parameters.
%
\item  The apparently-searchable star is a binary system, in which neither of 
the components is a searchable star, but at least one is of the desired type. 
$N_\star$ is over-counted by one for every such system.
\end{enumerate}

One might also imagine an apparently-searchable star which in no stellar 
component is of the desired type.
%For single stars, this may simply be from erroneous stellar parameters.
%In binaries, one might imagine the combined light of two stars somehow making 
%them resemble a single star of the desired type.  
We will not consider errors of this type.  We will assume that all the 
apparently-searchable stars are either single stars of
the desired type, or binaries in which either the primary or else both 
components are of the desired type.

@@

Errors (1)-(3) in calculating $N_{\rm det}$ can be broadly described as 
``radius misclassification'', while error (4) is an error in the assumed 
survey completeness.
%Note that in the above list, a planet cannot be both ``of desired type'' and 
%``orbiting a star of undesired type''. We ignore this type of error for 
%analytic convenience, but \texttt{reintroduce it later during numerics}.

Our approach is step-by-step starting from a very simple scenario.
In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
desired types of stars and planets.  
We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
of the true to apparent occurrence rate, taking into account all applicable 
errors.
In Sec.~\ref{sec:model_2} {\bf we allow for a distribution of the mass ratio 
$dn/dq \propto q^\gamma$} {\bf (not done yet)}, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
numerically for the case of planets orbiting Sun-like dwarf stars.
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
In Sec.~\ref{sec:model_3}, 
we assume the planets have a power-law radius distribution, ${\rm 
d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
all relevant errors.
Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between.
One consequence is that the claimed HJ occurrence rate discrepancy between 
radial velocity \& transit surveys gets much weaker when accounting for 
binarity.
In Sec.~\ref{sec:discussion} we summarize the models and discuss how they can 
be used to interpret transit surveys.

\begin{comment}
We then dig in, w/ numerics, to what happens with different assumed
completeness vs Rp functions.

This is b/c the more believable occ rate papers do not heavily rely on their
completeness calibration -- they know they're complete.
This is a much more understandable regime than the wonky completeness estimate 
dependent zone of eta-Earth for Kepler.

Finally, for numerics, we simulate a sample of 20,000 
apparently-searchable-stars (or however many "stars" appear to be searchable
for Earthlike planets in the Kepler field).  Basically this means a magnitude
limited survey of 20,000 ``stars'' that are either single or are binaries with
the primary of the desired type.  The sample has a realistic distribution of
binaries (occurrence and luminosity ratio).  We calculate 
eta(apparent)/eta(true).  We can also play with errors of type 4 in this
numerical model (i.e. we let the relative occurrence rates between primaries 
and secondaries vary).  (It may not worth doing so for the analytic 
work-- though it's pretty simple. The way to play with this is just present 
both limits, and then also present a few weight factors (Furlan+17 do a good 
job).

We discuss ``how do we interpret Fulton+ 2017?''
That paper did (or claims to have done) a good job at weeding out numerator
errors. If true, means $\Gamma(R_{\rm p})$ has the right shape, and just the 
wrong normalization. Important if we ever manage to convince ourselves 
``brightest transit hosts'' could work.

In passing, at the end, we mention how this changes attempts at eta-Earth 
measurements.
\end{comment}

\begin{comment}
UNSOLVED PROBLEMS W/ THE NUMERICAL MODEL-
* currently have not implemented self-consistent (L,M,R) relation between
primaries and secondaries.
Wang et al (2015) avoided that problem by doing a similar empirical thing to
what I currently have implemented. But I do not think that was a good paper.
Ciardi et al (2015) have an approach that was Kepler-specific -- they used
the same Dartmouth isochrones used for KIC stellar parameters.
But they did not conserve $dn/dm_x$, for $m_x$ apparent mag in band x.
I think the latter choice might be better for Kepler-specific discussion.
But it's less-nice if we want this to be general enough for e.g., occ HJ
rates w/ TESS.

My main gripe w/ taking the 150k KIC stars \& adding binaries is that less
is known about those parameters and biases. Using e.g., Galaxia has the major
benefit of you understand what's been cooked in much better (it's minimal).

I think if we want to talk about Kepler, it's the best possible approach
though.

If we want to talk about Fulton+17 we can go further, and apply the CKS cuts.

Some questions based on the current draft:

- Why should we ever consider $F_{lim} < F_{min}$?

LB: in practice, surveys look at many stars for which they could not detect
Rp=Re planets, but could detect Rp=10Re planets.  One possible error is
miscounting which stars are which (i.e. getting completeness corrections
wrong). So this was more for plausbility than b/c it's immediately needed in
the analytics. I can cut it.

- Why do we need to consider the distribution of S/N, as opposed to simply the
limiting distance for detection?

LB: for "model 1" (fixed stars, fixed planets, fixed light ratio binaries)
you're correct, it's not immediately necessary.

For this model, I did it as a way to think through the completeness effects
systematically -- i.e. how for the twin-binary case, you get a completeness
fraction of 1/8, though your limiting distance is sqrt(2) times bigger. (Even
when your completeness for the singles is 100%).

If we go the path of completing "model 2"'s analytics (vary light ratio, fixed
primary luminosity, fixed planet), I think this formulation would become
necessary in order to actually evaluate the completeness fractions.

(I never completed these analytics b/c of the L(M) mess + confusion about
assumed stellar masses+radii. Power law L(M) helps this... I think there needs
to be an extra assumption about the stellar mass+radius change though that we
have not spelled out)

- I don't think the bandpass, effective area, photon count rate, etc.  are
needed, either. All that matters is the limiting distance.

Are needed for what? 
In Model 1, for analytic expressions of the occurrence rate, I agree.
For analytic expressions of the number of detected planets, I disagree.

For Model 2, I'm not sure because I got bogged down. I suspect that for
occurrence rate errors, you are correct.

The analytic expression I gave for the occurrence rate depends on this "xi"
factor, which needs to incorporate an assumption about how Astronomer A
misestimates the completeness (relatedly, how he misestimates Rstar).  I never
specified that assumption.

- In Eq. 4, is this a traditional definition of BF?  I would have defined it as
nd/(ns+nd) or perhaps 2*nd/(ns+ 2*nd).  Let's make sure we are using
traditional terminology, and also interpreting the literature correctly, i.e.,
is the right number really 0.45?

Oops. nd/(ns+nd) = BF = MF (if assuming all multiple systems are binaries).
This minor change propagates.

Touching up on correct terminology, from Duchene \& Kraus 2013, Sec 2.1:

"Frequency of multiple systems" (MF)
vs
"Companion frequency" (CF; average \# of companions per target, can exceed 
100\%)

my understanding of the terms:

survey 1: {*, **, *, ***, *, *}

6 systems.
2 are multiple.
MF = 2/6 = 1/3
CF = (0+1+0+2+0+0)/6 = 3/6 = 1/2

survey 2: {**, **, ***, ****, *}

5 systems
4 are multiple
MF = 4/5
CF = (1+1+2+3+0)/5 = 6/5

DK2013 Sec 3.1.1 give (citing Raghavan+ 2010)
MF^{MS}_{1-1.3Msun} = 50 \pm 4%
MF^{MS}_{0.7-1Msun} = 41 \pm 3%

They also give
CF^{MS}_{1-1.3Msun} = 75 \pm 5%
CF^{MS}_{0.7-1Msun} = 56 \pm 4%

I understand the first pair of numbers to mean:

In a volume-limited survey (of the stellar neighborhood), if you select systems
for which the primary is from 1-1.3Msun, 50% of them will be multiple.

I take the second pair of numbers to mean:

In a volume-limited survey (of the stellar neighborhood), if you select systems
for which the primary is from 1-1.3Msun, each system will on average have 0.75
companions.

So YES, need to correct this throughout.
\end{comment}

\section{Model \#1: fixed stars, fixed planets, fixed light-ratio binaries}
\label{sec:model_1}

Consider the following idealized survey.
An astronomer observes the entire sky for some duration with a detector of 
known area and bandpass.
The detector is photon-noise limited.
They are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. %For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
They are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. %For instance, G2V dwarfs.
They detect the desired type of planet when they observe signals 
above some signal-to-noise floor.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux required for detection, $F_{\rm min}$.
Therefore they observe all of the ``points'' on-sky with apparent 
magnitude $m < m_{\rm min}$, or equivalently with energy flux in their 
bandpass greater than the limiting flux $F > F_{\rm min}$.
After detecting all planets with ${\rm S/N} > {\rm (S/N)_{min}}$, they now 
wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.

Assume this astronomer lives in a universe in which the following assumptions 
hold.
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twins.
	The ratio of the number density of binary systems to all systems in a 
	volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows.
    In single star systems, a fraction $\Gamma_{t,s}$ of stars 
	have planets of radius $R_p$, with orbital period $P$.
	In double star systems, a fraction $\Gamma_{t,d}$ of primary stars have 
	planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets 
	at equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
\end{itemize}

The assumption of [Pepper et al, 2003] in a very similar context was that the 
observer can correctly pre-select all of the ``searchable stars''.
This simplification would allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

However, any signal-to-noise limited survey of a population that includes 
binaries cannot ignore incompleteness.
%If we set the limiting flux to give 100\% completeness for single stars, we 
%will unwittingly select binaries near that flux limit, for which dilution 
%will push transit signals below the detection threshold.
This is because any given flux limit maps onto three distinct ``maximum 
distances'' for a population of single stars and twin binaries.
First, there is $d_{{\rm max},s}$: the maximum distance out to which the 
desired type of planet is detectable around single stars. Typically this is 
set to be the distance out to which stars in a SNR-limited survey are 
selected, since these are the ``searchable stars''.
Using the same minimum flux as that defining $d_{{\rm max},s}$, double stars 
are then selected out to a larger limiting distance, $d_{{\rm 
max},d}$~--~effectively a Malmquist bias.

However, the distance out to which double stars are {\it selected} is 
different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
be {\it detected} (around the primary of a double star system).
For the binary twin population, all planet and stellar radii are identical, so 
${\rm S/N} \propto \mathcal{D} L^{1/2} d^{-1}$, and
\begin{equation}
d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
(1 + \gamma_R)^{-1} d_{{\rm max},d}.
\end{equation}
For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
are in fact binary systems can yield detectable planets (the ratio of 
volume in which planets can be detected to the actual selected volume).

\begin{comment}
We note in passing that we are considering a SNR-limited survey, which in this 
case is the same as a magnitude-limited one.
The observer might prefer to perform a completeness-limited survey, which in 
this case would be the same as a volume-limited one with limiting distance
$d_{{\rm max},d}^{\rm p}$.
This change might circumvent various biases.
\end{comment}
With our model defined, we ask:

\subsection{What is the true occurrence rate?}
\label{sec:true_rate}

The ``true occurrence rate'' $\Gamma_t$ is the average number of planets of 
the desired type per star of the desired type:
\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + (1+w_{d2}) \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}
where $N_s$ is the number of single star systems, and $N_d$ the number of 
double star systems. 
The factor of 2 in the total number of stars accounts for the fact that there 
are twice as many stars in double star systems.

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
These latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
multiple systems that are higher order multiples as doubles.

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
stated in terms of a limiting magnitude $m_{{\rm min}}$.
In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
consequence of imposing a magnitude cut, the maximum distance to which binary 
stars will be selected is greater than that of single stars.
The ratio of double to single systems in the observed sample is 
then
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s}
	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
\label{eq:beta_model1}
\end{align}
In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
SNR-limited sample.
Since the completeness fraction for binary systems is $1/8$, there are 
$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
planets as single star systems in the system that can yield planets.

\subsection{What is the measured occurrence rate?}
% % % % % % % % % % % %

The observer who has never heard of binary star systems wants to find the 
occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars of 
radius $R_1$ and luminosity $L_1$.
They detected $N_{\rm det}$ transit signals that appear to be from planets of 
this desired type.
The occurrence rate they then find is
\begin{equation}
\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
\end{equation}

Since only the transit depth is observed, the planetary radii apparent to this 
astronomer, $R_{\rm p, a}$, are different from the true planetary radii 
$R_{\rm p, t}$. 
This is because the assumed stellar radii $R_{\star,{\rm a}}$ may differ from 
the true stellar radii $R_{\star,{\rm t}}$, and the flux will be diluted:
\begin{equation}
R_{p,{\rm a}} = R_{p,{\rm t}} 
\frac{R_{\star,{\rm a}}}
		 {R_{\star,{\rm t}}}
		 \,\mathcal{D}^{1/2},
\end{equation}
where the dilution parameter $\mathcal{D}$ is defined for binary systems as
\begin{align}
\mathcal{D} &= 
\Bigg\{\begin{array}{lr}
L_1 / L_d, & \text{if\ planet\ orbits\ primary}\\
\gamma_R L_1 / L_d, & \text{\ if\ planet\ orbits\ secondary}
\end{array} \nonumber\\
&=
\Bigg\{\begin{array}{lr}
(1+\gamma_R)^{-1}, & \text{if\ planet\ orbits\ primary}\\
(1 + \gamma_R^{-1})^{-1}, & \text{if\ planet\ orbits\ secondary}
\end{array} 
\label{eq:dilution}
\end{align}
where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.

In our twin binary thought-experiment, dilution means that all planets 
detected in binary systems will have apparent radii of $R_p/\sqrt{2}$.
Since for twin binaries the stellar radii are the same, the transit 
probability will be correctly calculated.

The total number of planet detections, irrespective of whether they are ``of 
the desired type'', is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
Expressing each individually,
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
and $f_{s,c}$ is the fraction of these transiting planets that are observed 
with signal to noise greater than the minimum detection threshold (the 
completeness). 
For our SNR-limited, twin-binary survey, $f_{s,g} = R_1/a\equiv f_g$, and 
$f_{s,c}=1$.

Analogously,
\begin{equation}
N_{{\rm det},d} = (1+w_{d2}) N_d \Gamma_{t,d} f_{d,g} f_{d,c},
\label{eq:N_det_d}
\end{equation}
where now $(1+w_{d2}) N_d \Gamma_{t,d}$ is the number of planets in the double 
star systems of the sample, the geometric transit probability $f_{d,g}=R_1/a$ 
for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
presented in the initial discussion of limiting distances.

The occurrence rate this astronomer derives for planets of assumed 
radius $R_p$ orbiting stars thought to be of the desired type is then
\begin{equation}
\Gamma_{\rm a} = \frac{\Gamma_{t,s}N_s}{N_s + N_d}.
\end{equation}

\subsection{Errors on derived occurrence rates for twin-binary thought 
experiment}
The ratio of the true to calculated occurrence rates for planets of radius 
$R_p$, $X_\Gamma$, is
\begin{align}
X_\Gamma &=
\frac{\Gamma_t}{\Gamma_{\rm a}} \\
&=
\frac{\Gamma_{t,s}N_s + (1+w_{d2})\Gamma_{t,d}N_d}{\Gamma_{t,s} N_s } \cdot 
\frac{N_s + N_d}{N_s + 2N_d} \\
&= 
\left( 
1 + \frac{(1+w_{d2}) \Gamma_{t,d}}{\Gamma_{t,s}} \beta
\right) \cdot 
\frac{1+\beta}{1+2\beta},
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = \frac{\rm BF}{1- {\rm BF}} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.44$, if we assume that all stars 
have equal occurrence rates, $\Gamma_{t,s} = \Gamma_{t,d}$ and 
$w_{d2}=1$, then we find $X_\Gamma = 1+\beta = 1+ 2^{3/2}\approx4$~--~ the 
observed occurrence rate is underestimated by a factor of 4.

For the planets around double stars with observed radii $R_p 
\sqrt{\mathcal{D}}$, there is also the error of their misclassification.
The fraction of detections with misclassified radii can be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\zeta},
\end{equation}
for
\begin{equation}
\zeta \equiv \frac{N_s \Gamma_{t,s} f_{s,g} f_{s,c}}{
	 (1+w_{d2})N_d \Gamma_{t,d} f_{d,g} f_{d,c }}.
\end{equation}

For our nominal G2V dwarf case, the twin binaries with equal 
occurrence rates gives $\zeta = 4/\beta = \sqrt{2}$, so 41\% of detections 
have misclassified radii, all of $R_p/\sqrt{2}$.

Of course more generally, the observer might define their 
desired ``type of planet'' to be anything within some margin of $R_p$.
Considering this for the idealized twin-binary case, if the margin includes 
the planets in binary systems, the number of detections ``of the desired 
type'' would change.
The correction factor $X_\Gamma$ would then become
\begin{equation}
X_\Gamma = 
\frac{\Gamma_{t,s} + (1+w_{d2})\Gamma_{t,d}\beta}{\Gamma_{t,s} + 
(1+w_{d2})\Gamma_{t,d} f_{d,c} } \cdot 
\frac{1+\beta}{1+2\beta},
\end{equation}
which simplifies to $(1+2f_{d,c})^{-1} (1+\beta) = 4(1+\beta)/5\approx 3$ for 
our twin-binary thought experiment.
This makes sense: increasing the radius interval of ``desired planets'' 
decreases the error in the numerator, but the error of mis-counted stars in 
the denominator remains identical.

\section{Model \#2: fixed primaries, fixed planets, varying light-ratio 
secondaries}
\label{sec:model_2}

The model of Sec.~\ref{sec:model_1} is idealized, but it helps
develop intuition.
Now, we let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
of star systems.
It does so because the underlying mass ratio $q=M_2/M_1$ varies.
We keep the primary mass fixed as $M_1$, which is also the mass of all single 
stars.

>>>>

@@

An astronomer who does not believe in stellar multiplicity wants to measure the
mean number of planets of a certain type per star of a certain type.
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that seem to be of the desired type, 
and searchable for the planet.  ``Searchable stars'' are the stars for which 
planets are observed with 100\% detection efficiency. 
Correcting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Lambda_{\rm a}$:

\begin{equation}
\Lambda_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
\end{equation}

@@

Here we focus on problems that arise from the fact that stars of the desired
type often exist in multiple star systems.
For simplicity, we only consider binaries, and we assume that they are all 
spatially unresolved.

An immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~the occurrence rate of planets may differ 
between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval of say $\Delta 
M_\star$, it must implicitly marginalize over stellar multiplicity.
Otherwise, one must discuss ``occurrence rates in single star systems'', 
``occurrence rates about primaries'', and
``occurrence rates about secondaries'' (this has been noted by Wang et al., 
2015)

Outside of astrophysical differences, there are observational biases.
A given apparently-searchable star may truly be a single star of the desired 
type. If not one of the following must be true:
\begin{enumerate}	
    \item The apparently-searchable star is a binary system, with two 
    searchable 
    stars of the desired type. $N_\star$ is under-counted by one for every 
    such 
    system.
    %
    \item The apparently-searchable star is a binary system, with one 
    searchable star of the desired type. 
    $N_\star$ is correctly counted for every such system.
    There may be systematic errors in stellar parameter estimates of such 
    systems, but in this work we neglect these.
    %
    \item  The apparently-searchable star is a binary system, in which neither 
    of the components is a searchable star, but at least one is of the desired 
    type. 
    $N_\star$ is over-counted by one for every such system.
\end{enumerate}

One might also imagine an apparently-searchable star which in no stellar 
component is of the desired type.
We will not consider errors of this type.  We will assume that all the 
apparently-searchable stars are either single stars of
the desired type, or binaries in which either the primary or else both 
components are of the desired type.

@@

Errors (1)-(3) in calculating $N_{\rm det}$ can be broadly described as 
``radius misclassification'', while error (4) is an error in the assumed 
survey completeness.

\paragraph{The Hot Jupiter Rate Discrepancy}

Radial velocity surveys of nearby bright Sun-like stars have reported hot 
Jupiter ($a<0.1\,{\rm AU}, P\lesssim 10\,{\rm day}$) occurrence rates of
$12\pm 1$, $15\pm 6$, $8.9 \pm 3.6$, and $12.0 \pm 3.8$ per thousand stars
(Marcy et al 2005, Cumming et al 2008, Mayor et al 2011, Wright et al 2012 
respectively).

In transit surveys, OGLE-III reported an absolute rate of $3.1^{+ 
4.3}_{-1.8}$ hot Jupiters with $P<5\,{\rm days}$ per thousand stars (Gould et 
al 2006).
The SuperLupus survey quoted $10^{+27}_{-8}$ per thousand stars, allowing 
planets with periods less than 10 days (Bayliss \& Sackett 2011).
Through the {\it Kepler}\ survey, Howard et al. (2012) later reported a rate 
of 
$4 \pm 1$ per thousand Sun-like stars.
This rate was for $P<10\,{\rm day}, 8-32r_\oplus$ planets, and 
was specific to their ``solar subset''\footnote{The ``solar subset'' was 
defined as $4100<T_{\rm eff}/{\rm K}<6100$, {\it Kepler}\ 
magnitude $<15$, $4.0 < \log g < 4.9$, and only took planets with measured 
signal to noise $>10$.
}.
Expanding their sample to fainter magnitudes, they quoted a rate of $5 \pm 
1$.
Expanding down to $r_p>5.6r_\oplus$, to avoid excluding hot Jupiters reported 
by RV surveys, they reported $7.6 \pm 1.3$.
Recent work by Petigura et al. using the updated parameters of the 
California-{\it Kepler}\ Survey has found a rate of $X.X \pm 
Y.Y$ (2017, in preparation)

The trend is that hot Jupiter occurrence rates measured by transit 
surveys seem to be marginally lower than those found by radial velocity 
surveys.
While the actual discrepancy is of sub-$3\sigma$ significance,
one reason to expect a difference in the rates is that the populations have 
different metallicities.
As originally argued by Gould et al. (2006), the RV sample is biased towards 
metal-rich stars, which have been measured by RV surveys to preferentially 
host more giant planets (Santos et al 2004, Fischer and Valenti 2005).
The {\it Kepler}\ sample specifically has been measured to be more metal poor 
than the local neighborhood, with a mean metallicity of $[{\rm M/H}]_{\rm 
mean}\approx -0.05$ (Dong et al., 2014; Guo et al., 2017).
Studying the problem in detail, Guo et al. recently argued that the 
metallicity difference could account for a $\approx 0.1\%$ difference in the 
measured rates between the CKS and {\it Kepler}\ samples~--~not a $\approx 
0.5\%$ difference.
These authors concluded that ``other factors, such as binary contamination and 
imperfect stellar properties'' must also be at play (Guo et al., 2017).

Aside from surveying stars of different metallicities, radial velocity and 
transit surveys differ in how they treat binarity.
Radial velocity surveys typically reject both visual and spectroscopic binaries
({\it e.g.}, Wright et al. 2012).
Transit surveys typically observe binaries, but the question of whether they 
were searchable to begin with is usually left for later interpretation.
In spectroscopic follow-up of transiting candidates, the prevalence of 
astrophysical false-positives may also lead to a bias against confirmation of 
transiting planets in binary systems.

Ignoring these complications, in this work we focus on whether
binarity might intrinsically bias occurrence rate measurements, simply 
through its effects on the number of searchable stars and the apparent radii 
of detected planets.

To progressively gain intuition, we study the following idealized transit 
surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed planets and primaries, varying secondaries;
    \item Model \#3: fixed primaries, varying planets and secondaries.
\end{itemize}
In Sec.~\ref{sec:numerical_methods}, we introduce our numerical approach 
to the problem, and in Sec.~\ref{sec:analytic_preliminaries} we clarify our 
terminology.
We present the analytic and numerical results in 
Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where each subsection corresponds 
to each model above.
We interpret these calculations throughout, and in 
Sec.~\ref{sec:discussion} discuss their relevance to various questions in 
the interpretation of transit survey occurrence rates.
